---
title: "ds202_lab3"
author: 'Chase Weida; Github: cweida'
date: "2/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(ggplot2)
library(ggpubr)
```

### 1
How does the happiness of a respondent relate to the marriage status?

```{r}
dat <- read_xls('../ds202_lab3/Data/GSS.xls')
colnames(dat) <- c("Year", "id", "Work_per_wk", "Marital_Status", "Spouse_Education", "Sex", "Political_Party", "General_Happiness", "Ballot_used")
ggplot(dat, aes(x = General_Happiness, fill = Marital_Status)) + geom_bar()
```

### 2
Does the sex of the respondent affect the relationship you found in Q1?

```{r}
plotMale <- ggplot(dat) + ggtitle('Happiness vs Marital Status in Females') + geom_bar(data = dat[dat$Sex == 'Female', ], aes(x = General_Happiness, fill = Marital_Status)) + theme(legend.position = "none")  
plotFemale <- ggplot(dat) + ggtitle('Happiness vs Marital Status in Males') + geom_bar(data = dat[dat$Sex == 'Male', ], aes(x = General_Happiness, fill = Marital_Status))
ggarrange(plotMale, plotFemale, common.legend = TRUE)

```

### 3
How does the happiness of a respondent relate to the political party affiliation?

```{r}
dat1 <- dat[!is.na(dat$General_Happiness),]
dat1$Political_Party <- factor(dat1$Political_Party, levels=c("No answer", "Don't know", "Other party", "Strong democrat", "Not str democrat", "Ind,near dem", "Independent", "Ind,near rep", "Not str republican", "Strong republican"))
ggplot(dat1, aes(x = Political_Party, fill = General_Happiness))+ geom_bar() + coord_flip()
```

Democrats and Democrat leaning independents tended to be happier than their Republican counterparts.

### 4a
How does the poltical party affiliation relate to the year the survey was conducted?

```{r}
dat2 <- dat[!(dat$Political_Party == "Don't know" | dat$Political_Party == "No answer"),]
dat2$Political_Party <- factor(dat2$Political_Party, levels=c("No answer", "Don't know", "Other party", "Strong democrat", "Not str democrat", "Ind,near dem", "Independent", "Ind,near rep", "Not str republican", "Strong republican"))
ggplot(dat2, aes(x = Year, fill = Political_Party)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent)
```

There was a rise in the number of Republicans and Republican leaning independents between 1980 and 1990 but then it went back down.

### 4b
What is the relation between marital status and political party affiliation?

```{r}
dat3 <- dat[!(dat$Political_Party == "Don't know" | dat$Political_Party == "No answer" | is.na(dat$Political_Party)),]
dat3 <- dat3[!(is.na(dat3$Marital_Status) | dat3$Marital_Status == "No answer"),]
dat3$Political_Party <- factor(dat3$Political_Party, levels=c("No answer", "Don't know", "Other party", "Strong democrat", "Not str democrat", "Ind,near dem", "Independent", "Ind,near rep", "Not str republican", "Strong republican"))
ggplot(dat3, aes(x = Marital_Status, fill = Political_Party)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent)
```

It is more likely for a married or widowed person to be a Republican or a Republican leaning independent, than it is for a divorced, separated, or never married individual.
